Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-09-28 ejap.manuscript.RMD
ejap format

---
title: ""
author: ""
bibliography: resources/ribose-paper.bib
csl: resources/ejap.csl
output:
  word_document:
    reference_docx: resources/reference-template-02.docx
editor_options: 
  chunk_output_type: console
---

@@

#### Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

@@

\pagebreak
<div custom-style="abstract_style">
# ABSTRACT

@@

*Purpose:* Muscle hypertrophy, induced by resistance training, is determined by ribosomal biogenesis. High-glucose treatment may augment the transcription of ribosomal RNA, as evident from in vitro studies. However, its effect on resistance training-induced adaptations remains unexplored.

*Methods:* Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition. Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension). Micro biopsies were sampled from m. vastus lateralis before and after the intervention.

*Results:* Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of cellular myelocutmatosis oncogene, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`). There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).

*Conclusion:* Glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.

*Keywords:* Glucose; hypertrophy; resistance training; ribosome; skeletal muscle

```{r abbreviation table}
### Abbreviations table

@@

## Flextable
abb.tab <- tab.dat %>%
  flextable() %>%
  set_caption(caption = "Abbreviations",
              align_with_table = FALSE) %>%
  set_table_properties(layout = "autofit")

@@

@@

# Statements and Declarations
This study was financed through research funds from the Inland Norway University of Applied Sciences. There were no economical vested interest and everyone participating as researchers and associates, received only a regular salary during this study.

## AUTHOR CONTRIBUTIONS

## Acknowledgements
Gratitude are in order to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work in the training intervention. The authors would also like to express their gratitude to all the participants, staying commited to this demanding protocol in an impressive way. We also thank Håvard Nygaard for lending your expertise and technical skills to improve this project.

\pagebreak
# INTRODUCTION

@@

# MATERIALS AND METHODS

@@

## PARTICIPANTS

@@

## EXPERIMENTAL DESIGN

@@

## DIETARY INTERVENTION

@@

## ASSESSMENT OF MUSCULAR STRENGTH

@@

## RESISTANCE TRAINING PROTOCOL

@@

## SAMPLING OF MUSCLE TISSUE AND BLOOD

@@

## TOTAL RNA EXTRACTION AND REAL-TIME REVERSE TRANSCRIPTION POLYMERASE CHAIN REACTION
```{r totrna-extraction}

@@

## PROTEIN EXTRACTION AND IMMUNOBLOTTING

@@

## STATISTICS AND DATA ANALYSIS

@@

# RESULTS

@@

## Markers of ribosome biogenesis

### Total RNA and ribosomal RNA

@@

### Protein

@@

# DISCUSSION

@@

### Conclusion

@@

## DATA AVAILABILITY

@@

## SUPPLEMENTAL MATERIAL

@@

# FIGURE LEGENDS

@@

# TABLES

```{r baseline table, echo = FALSE}

@@

  ## Make it to flextable
  flextable() %>%
  ## Change headers
  compose(i = 1, j = c(1, 2, 3, 4, 5, 6, 7, 8, 9), part = "header", 
          value = c(as_paragraph("Sex"),
                    as_paragraph("n"), 
                    as_paragraph("Age (yrs)"), 
                    as_paragraph("Stature (cm)"), 
                    as_paragraph("Body mass (kg)"),
                    as_paragraph("Lean mass (kg)"),
                    as_paragraph("60\U00BA sec", as_sup("-1")), 
                    as_paragraph("240\U00BA sec", as_sup("-1")), 
                    as_paragraph("0\U00BA sec", as_sup("-1")))) %>%
  add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3)) %>%
  set_caption(caption = "Table 1. Participant characteristics",
              align_with_table = FALSE)

@@

```
Values are means ± SD. 60º sec-1 = 60 degrees per second movement speed (isokinetic), 240º sec-1 = 240 degrees per second movement speed (isokinetic), 0º sec-1 = 0 degrees per second movement speed (isometric).

```{r primer table, echo = FALSE}

@@

## Flextable
primer.tab <- tab.dat %>%
  flextable() %>%
  compose(i = 1, j = c(1, 2, 3, 4), part = "header",
          value = c(as_paragraph("Gene"),
                    as_paragraph("Sequence (forward - reverse)"),
                    as_paragraph("Ct mean (SD)"),
                    as_paragraph("E"))) %>%
  set_caption(caption = "Table 2. Primer Sequences",
              align_with_table = FALSE) %>%
  set_table_properties(layout = "autofit")

@@

```
Values of CT are means ± SD. rRNA = ribosomal RNA, E = primer efficiency. Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.

>>>>

---
title: ""
author: ""
bibliography: resources/ribose-paper.bib
csl: resources/ejap.csl
output:
  word_document:
    reference_docx: resources/ejap-template.docx
editor_options: 
  chunk_output_type: console
---

@@

# Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

@@

\pagebreak
<div custom-style="abstract_style">
## Abstract

@@

**Purpose:** Muscle hypertrophy, induced by resistance training, is determined by ribosomal biogenesis. High-glucose treatment may augment the transcription of ribosomal RNA, as evident from in vitro studies. However, its effect on resistance training-induced adaptations remains unexplored.

**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition. Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension). Micro biopsies were sampled from m. vastus lateralis before and after the intervention.

**Results:** Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of cellular myelocutmatosis oncogene, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`). There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).

**Conclusion:** Glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.

**Keywords:** Glucose; hypertrophy; resistance training; ribosome; skeletal muscle

**Abbreviations**
```{r abbreviation table}
### Abbreviations table

@@

## Flextable
abb.tab <- tab.dat %>%
  flextable() %>%
  set_table_properties(layout = "autofit")

@@

### Statements and Declarations
This study was financed through research funds from the Inland Norway University of Applied Sciences. There were no economical vested interest and everyone participating as researchers and associates, received only a regular salary during this study.

### Author contributions

### Acknowledgements
Gratitude are in order to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work in the training intervention. The authors would also like to express their gratitude to all the participants, staying commited to this demanding protocol in an impressive way. We also thank Håvard Nygaard for lending your expertise and technical skills to improve this project.

@@

\pagebreak
## Introduction

@@

## Materials and methods

@@

### Participants

@@

### Experimental design

@@

### Dietary intervention

@@

### Assessment of muscular strength

@@

### Resistance training protocol

@@

### Sampling of muscle tissue and blood

@@

### Total RNA extraction and real-time reverse transcription polymerase chain reaction
```{r totrna-extraction}

@@

### Protein extraction and immunoblotting

@@

### Statistics and data analysis

@@

## Results

@@

### Markers of ribosome biogenesis

*Total RNA and ribosomal RNA*

@@

*Protein*

@@

## Discussion

@@

*Conclusion*

@@

### DATA AVAILABILITY

@@

### SUPPLEMENt MATERIAL

@@

# Figure/table legends

@@

# Tables

**Table 1. Participant characteristics:** Values are means ± SD. 60º sec-1 = 60 degrees per second movement speed (isokinetic), 240º sec-1 = 240 degrees per second movement speed (isokinetic), 0º sec-1 = 0 degrees per second movement speed (isometric).
```{r baseline table, echo = FALSE}

@@

  ## Make it to flextable
  flextable() %>%
  ## Change headers
  compose(i = 1, j = c(1, 2, 3, 4, 5, 6, 7, 8, 9), part = "header", 
          value = c(as_paragraph("Sex"),
                    as_paragraph("n"), 
                    as_paragraph("Age (yrs)"), 
                    as_paragraph("Stature (cm)"), 
                    as_paragraph("Body mass (kg)"),
                    as_paragraph("Lean mass (kg)"),
                    as_paragraph("60\U00BA sec", as_sup("-1")), 
                    as_paragraph("240\U00BA sec", as_sup("-1")), 
                    as_paragraph("0\U00BA sec", as_sup("-1")))) %>%
  add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3))

@@

**Table 2. Primer sequences:** Values of Ct are means ± SD. rRNA = ribosomal RNA, E = primer efficiency. Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions. 
```{r primer table, echo = FALSE}

@@

## Flextable
primer.tab <- tab.dat %>%
  flextable() %>%
  compose(i = 1, j = c(1, 2, 3, 4), part = "header",
          value = c(as_paragraph("Gene"),
                    as_paragraph("Sequence (forward - reverse)"),
                    as_paragraph("Ct mean (SD)"),
                    as_paragraph("E"))) %>%
  set_table_properties(layout = "autofit")

@@
